---
title: "SpotifyAPI (from SpotifyR Workshop at R User Group Finland)"
author: "Marc Eixarch (orig.), modified by Antti Rask"
date: "2023-03-30"
output: html_document
---

Testing the Spotify API using the [{spotifyr}](https://github.com/charlie86/spotifyr) package created by Charlie Thompson and currently maintained by Daniel Antal.

EDIT: This version only includes the part that creates a new Spotify playlist.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r, warning=FALSE}
library(conflicted)
library(spotifyr)
library(tidyverse)

source("secret.R")
```

## Getting the SpotifyAPI Access Token

```{r}
Sys.setenv(SPOTIFY_CLIENT_ID     = .SPOTIFY_CLIENT_ID)
Sys.setenv(SPOTIFY_CLIENT_SECRET = .SPOTIFY_CLIENT_SECRET)

access_token <- get_spotify_access_token()
```

## Creating the function for getting the authorization code

```{r}
get_authorized <- function(.scope) {
    
    get_spotify_authorization_code(
        client_id     = .SPOTIFY_CLIENT_ID,
        client_secret = .SPOTIFY_CLIENT_SECRET,
        scope         = .scope
    )
}
```

## Creating customised recommended Playlists

### Finding out my top artists

```{r}
my_top_artists <- get_my_top_artists_or_tracks(
    type          = "artists", 
    limit         = 50,            # the maximum
    time_range    = "medium_term", # about 6 months history,
    authorization = get_authorized("user-top-read")
)

my_top_artists %>%
    select(name, genres, popularity) %>% 
    arrange(desc(popularity))
```

## Creating a playlist

### Get recommendations based on the chosen attributes and artists

```{r}
my_happy_and_energetic_songs <- get_recommendations(
    seed_artists = head(my_top_artists, 5) %>% pull(id),
    min_energy   = 0.6,
    min_valence  = 0.6
)
```

### Create an empty playlist

```{r}
.playlist_id <- create_playlist(
    user_id       = .USER_ID,
    name          = str_glue("My Happy and Energetic Songs ({today()})"),
    description   = "Generated with R!",
    authorization = get_authorized("playlist-modify-public")
)$id
```

### Populate the created playlist

```{r}
add_tracks_to_playlist(
  playlist_id   = .playlist_id,
  uris          = my_happy_and_energetic_songs$id,
  authorization = get_authorized("playlist-modify-public")
)
```
