---
title: "SpotifyAPI (from SpotifyR Workshop at R User Group Finland)"
author: "Marc Eixarch (orig.), modified by Antti Rask"
date: "2023-03-30"
output: html_document
---

Testing the Spotify API using the [{spotifyr}](https://github.com/charlie86/spotifyr) package created by Charlie Thompson and currently maintained by Daniel Antal.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r, warning=FALSE}
library(conflicted)
    conflicts_prefer(dplyr::filter)
library(extrafont)
library(ggExtra)
library(ggrepel)
library(spotifyr)
library(tidyverse)

source("secret.R")
```

## Getting the SpotifyAPI Access Token

```{r}
Sys.setenv(SPOTIFY_CLIENT_ID     = .SPOTIFY_CLIENT_ID)
Sys.setenv(SPOTIFY_CLIENT_SECRET = .SPOTIFY_CLIENT_SECRET)

access_token <- get_spotify_access_token()
```

## Creating the function for getting the authorization code

```{r}
get_authorized <- function(.scope) {
    
    get_spotify_authorization_code(
        client_id     = .SPOTIFY_CLIENT_ID,
        client_secret = .SPOTIFY_CLIENT_SECRET,
        scope         = .scope
    )
}
```


## Part 1

### Getting my playlists

```{r}
my_playlists <- get_my_playlists(
    limit         = 50,
    authorization = get_authorized("playlist-read-private")
) %>%
    as_tibble()

my_playlists %>% 
    select(name, owner.display_name, tracks.total)
```

### Checking categories

```{r}
get_categories() %>% pull(name)
```

### Getting my top artists

```{r}
my_artists <- get_my_top_artists_or_tracks(
    limit = 2,
    authorization = get_authorized("user-top-read")
) %>% 
    pull(name)

my_artists
```

### Getting the audio features for the top artists

```{r}
my_artists_track_features <- bind_rows(
    map(
        my_artists,
        get_artist_audio_features,
        market = "US"
    )
) %>% 
    select(
        artist_id,
        artist_name,
        track_id,
        track_name,
        album_name,
        album_release_year,
        duration_ms,
        tempo,
        acousticness,
        danceability,
        energy,
        instrumentalness,
        liveness,
        loudness,
        speechiness,
        valence
    ) %>%
    as_tibble()

my_artists_track_features
```

### Getting summary statistics

```{r}
my_album_summary_stats <- my_artists_track_features %>%
    summarise(
        across(where(is.numeric), mean),
        tracks = n(),
        .by    = c(artist_name, album_release_year, album_name)
    )

my_album_summary_stats
```

### Live Albums

#### Checking for Live Albums

```{r}
my_album_summary_stats %>% 
    filter(liveness > 0.29) %>% 
    select(artist_name, album_release_year, album_name, liveness)
```

#### Getting rid of the Live Albums

```{r}
my_artists_track_features_without_live <- my_artists_track_features %>%
    filter(liveness < 0.29)

my_album_summary_stats_without_live    <- my_album_summary_stats %>% 
    filter(liveness < 0.29)

my_artists_track_features_without_live
my_album_summary_stats_without_live
```

### Special Editions

#### Checking for Special Editions

```{r}
my_album_summary_stats_without_live %>%
    filter(str_detect(tolower(album_name), "edition") == TRUE)
```

#### Getting rid of the Special Editions

```{r}
my_artists_track_features_without_special <-my_artists_track_features_without_live %>%
    filter(str_detect(tolower(album_name), "edition") == FALSE)

my_album_summary_stats_without_special <- my_album_summary_stats_without_live %>%
    filter(str_detect(tolower(album_name), "edition") == FALSE)

my_artists_track_features_without_special
my_album_summary_stats_without_special
```

### Insights

#### Summary of the average album Energy

```{r}
my_album_summary_stats_without_special %>%
    ggplot(aes(album_release_year, energy, color = artist_name)) + 
    geom_line() +
    geom_point() +
    theme_classic()
```

#### Summary of all the features

```{r}
my_album_summary_stats_without_special %>%
    mutate(
        album_number = row_number(album_release_year),
        .by = artist_name
    ) %>%
    pivot_longer(
        cols      = !c(artist_name, album_release_year, album_name, album_number),
        names_to  = "feature",
        values_to = "score"
    ) %>% 
    filter(
        feature %in% c(
            "acousticness",
            "danceability",
            "energy",
            "instrumentalness",
            "speechiness",
            "valence"
        )
    ) %>% 
    ggplot(aes(album_number, score, color = artist_name)) + 
    geom_line() +
    geom_point() +
    facet_wrap(vars(feature), ncol = 2) +
    theme_classic()
```

#### Energy vs. Valence

```{r}
p <- my_artists_track_features_without_special %>%
    ggplot(aes(energy, valence, color = artist_name)) +
    geom_point() +
    geom_hline(yintercept = 0.5, color = "grey", linetype = "dashed") +
    geom_vline(xintercept = 0.5, color = "grey", linetype = "dashed") +
    theme_classic() +
    theme(legend.position = "bottom")

ggMarginal(p, type = "density", groupColour = TRUE, groupFill = TRUE)
```

#### Top tracks, annotated

```{r}
top_tracks <- bind_rows(
    map(unique(my_artists_track_features$artist_id), get_artist_top_tracks)
)

top_tracks %>% 
    select(id, popularity) %>%
    right_join(
        my_artists_track_features,
        by = join_by(id == track_id)
    ) %>%
    mutate(
        rank_top_song = row_number(desc(popularity)),
        .by = artist_name
    ) %>%
    ggplot(aes(energy, valence, color = artist_name)) +
    geom_point() +
    geom_label_repel(
        aes(
            label = case_when(
                rank_top_song <= 3 ~ track_name,
                TRUE ~ NA_character_
            )
        )
    ) + 
    geom_hline(yintercept = 0.5,  color = "grey", linetype = "dashed") +
    geom_vline(xintercept = 0.5,  color = "grey", linetype = "dashed") +
    annotate("text", 0.25 / 2, 1, label = "Hopeful Ballads",  fontface = "italic") +
    annotate("text", 1.75 / 2, 1, label = "Vibrant Cheerful", fontface = "italic") +
    annotate("text", 1.75 / 2, 0, label = "Vibrant Enraged",  fontface = "italic") +
    annotate("text", 0.25 / 2, 0, label = "Sad Ballads",      fontface = "italic") +
    labs(
        x     = "Energy",
        y     = "Valence",
        color = "Artist",
        title = "Songs mood"
    ) +
    theme_classic()
```

### Spotify Theme

Based on the [Monokai theme](https://github.com/datarootsio/artyfarty)

#### Creating the Spotify Theme

```{r}
spotify_colors <- list(
    dark_green  = "#1DB954",
    light_green = "#1ed760",
    black       = "#191414",
    light_grey  = "#b3b3b3",
    white       = "#FFFFFF"
)

theme_spotify <- function(){
    color.background = spotify_colors$black
    color.grid.major = spotify_colors$white
    color.text       = spotify_colors$dark_green
    color.text.2     = spotify_colors$light_green
    color.axis       = spotify_colors$white
    
    theme_bw(base_size = 13, base_family = "Gotham") +
        
        theme(
            # Axis
            axis.line.x  = element_line(color = color.grid.major, linewidth = 1),
            axis.line.y  = element_line(color = color.grid.major, linewidth = 1),
            axis.text.x  = element_text(size  = rel(0.95), color = color.text),
            axis.text.y  = element_text(size  = rel(0.95), color = color.text),
            axis.ticks   = element_line(color = NA),
            axis.title.x = element_text(size  = rel(1), color = color.text, vjust = 0),
            axis.title.y = element_text(size  = rel(1), color = color.text, vjust = 1.25),
            
            # Legend
            legend.background = element_rect(fill  = color.background),
            legend.key        = element_rect(fill  = color.background, color = NA),
            legend.text       = element_text(size  = rel(0.8), color = color.text.2),
            legend.title      = element_text(color = color.text, face = "bold"),
            
            # Panel
            panel.background   = element_rect(
                fill  = color.background,
                color = color.background
            ),
            panel.border       = element_rect(color = color.background),
            panel.grid.major   = element_line(
                color     = color.grid.major,
                linewidth = 0.4,
                linetype  = 2
            ),
            panel.grid.major.x = element_blank(),
            panel.grid.minor   = element_blank(),
            
            # Plot
            plot.background = element_rect(fill = color.background, color = color.background),
            plot.title      = element_text(
                color = color.text,
                size  = rel(1.2),
                hjust = 0.5,
                face  = "bold"
            )
        )
}

monokai_palette <- c(
    "#F92672",
    "#FD971F",
    "#F4BF75",
    "#A6E22E",
    "#A1EFE4",
    "#66D9EF",
    "#AE81FF",
    "#CC6633"
)
```

#### Using the Spotify Theme

```{r}
my_artists_track_features %>% 
    pivot_longer(
        cols = !c(
            track_name,
            artist_name,
            album_release_year,
            album_name,
            track_id,
            artist_id
        ),
        names_to  = "feature",
        values_to = "score"
    ) %>% 
    filter(
        feature %in% c(
            "danceability",
            "energy",
            "valence"
        )
    ) %>%
    ggplot(aes(feature %>% str_to_title(), score, color = artist_name)) +
    geom_boxplot(
        fill      = spotify_colors$black,
        linewidth = 1
    ) +
    labs(
        color = "Artist",
        title = "Artists' Average Scores on Music Features",
        x     = NULL,
        y     = "Score"
    ) +
    scale_color_manual(values = monokai_palette) +
    theme_spotify()
```


## Part 2

### Creating customised recommended Playlists

#### Finding out my top artists

```{r}
my_top_artists <- get_my_top_artists_or_tracks(
    type          = "artists", 
    limit         = 50,            # the maximum
    time_range    = "medium_term", # about 6 months history,
    authorization = get_authorized("user-top-read")
)

my_top_artists %>%
    select(name, genres, popularity) %>% 
    arrange(desc(popularity))
```

#### Finding out my top genres

```{r}
my_top_artists %>%
    unnest(genres) %>% 
    summarise(
        artists = n(),
        .by     = genres
    ) %>% 
    arrange(desc(artists)) %>% 
    head(10) %>%
    mutate(
        genres = genres %>% 
            str_to_title() %>%
            as_factor() %>%
            fct_reorder2(genres, artists, .desc = FALSE)
    ) %>%
    arrange(artists) %>%
    ggplot(aes(genres, artists)) +
    geom_segment(
        aes(genres, 0, xend = genres, yend = artists),
        color = spotify_colors$light_green
    ) +
    geom_point(size = 3, color = spotify_colors$dark_green) +
    coord_flip() +
    labs(
        title = "My Top 10 Favorite Genres",
        x     = "Genre",
        y     = "Number of Artists"
    ) +
    theme_spotify() + 
    theme(
        axis.line.x      = element_blank(),
        axis.line.y      = element_blank(),
        panel.grid.major = element_blank(),
    )
```

#### How mainstream is my taste in music?

```{r}
my_top_artists %>%
    ggplot(aes(popularity)) + 
    geom_density(color = spotify_colors$black, fill = spotify_colors$dark_green) +
    coord_cartesian(xlim = c(0,100)) +
    labs(
        x = "Popularity",
        y = "Density"
    ) +
    theme_spotify()
```


## Part 3

### Creating a playlist

#### Get recommendations based on the chosen attributes and artists

```{r}
my_happy_and_energetic_songs <- get_recommendations(
    seed_artists = head(my_top_artists, 5) %>% pull(id),
    min_energy   = 0.6,
    min_valence  = 0.6
)
```

#### Create an empty playlist

```{r}
.playlist_id <- create_playlist(
    user_id       = .USER_ID,
    name          = str_glue("My Happy and Energetic Songs ({today()})"),
    description   = "Generated with R!",
    authorization = get_authorized("playlist-modify-public")
)$id
```

#### Populate the created playlist

```{r}
add_tracks_to_playlist(
  playlist_id   = .playlist_id,
  uris          = my_happy_and_energetic_songs$id,
  authorization = get_authorized("playlist-modify-public")
)
```
